Parsing Postfix log files

### 2008-11-30 doc/logparser-other-parsers-review.tex
Improved the introduction and the LMA section.

@@

Ten other parsers, starting with the only previously published research in
this area that the author is aware of, have been reviewed as part of the
background research for this project.  Each of the parsers was tested with
the \numberOFlogFILES{} test log files described in \sectionref{parser
efficiency}.  None of the parsers reviewed perform the kind of advanced
parsing and correlation required for this project; all except one (XXX IS
THIS ACCURATE\@?) are intended to perform a specific parsing and reporting
task, whereas \gls{PLP} is designed to be a generic parser, extracting data
and leaving generation of reports from the data to other programs.  Some
parsers save the data extracted to a data store, but the majority discard
all data once they have finished running, making historical analysis
impossible.  The parsers reviewed all produce a report of varying
complexity and detail, whereas this project's parser does not produce a
report at all; that responsibility is deferred to a separate program, to be
developed later.  The parser developed for this project is designed to
enable much more detailed log file analysis by providing a stable platform
for subsequent programs to develop upon.

It is important to compare and contrast newly developed programs,
algorithms, and parsers against those already available, to accurately
judge what improvements, if any, are delivered by the newcomers.  There are
not that many previously developed Postfix log file parsers --- indeed it
was quite difficult to find ten parsers to review for this project --- and
the functionality offered ranges from quite basic to much more mature,
depending on the needs of the creator.  The programs are not reviewed from
an independent viewpoint; the objective is to compare and contrast each
program with \parsername{}.  There are some important differences between
\parsername{} and the parsers reviewed here:

@@

@@

    \item \parsername{} does not provide any reports based on the data it
        gathers.  This is by design: the parser is responsible for parsing
        log files and extracting data; further processing of that data is a
        separate concern, deferred to another program.  Using an \gls{SQL}
        database simplifies the process of generating such reports
        (discussed in \sectionref{database as API}); some simple examples
        are given in \sectionref{motivation}.

@@

    \item A minor difference is that most parsers do not handle compressed
        files; both \parsername{} and Splunk handle them transparently,
        without user intervention; Sawmill and Lire can be configured to
        support compressed files, but Sawmill exhibits a dramatic slow down
        when thus configured.  Although this is a minor difference, support
        for reading compressed log files is quite helpful, as it
        dramatically reduces the disk space required to store old log
        files.

@@

@@

XXX IS THE BDB STRUCTURE DOCUMENTED\@?

XXX DESCRIBE THE \gls{CSV} FORMAT\@.

XXX DESCRIBE THE MYSQL SCHEMA\@.

IS THE SAME DATA STORED FOR ALL FORMATS\@?

@@

@@

@@

\gls{LMA} does not appear to deal with any of the complications discussed
in \sectionref{complications}.  It does not differentiate between different
types of rejections, so it is not suitable for the purposes of this
project; the data about which restriction caused the rejection is
discarded, whereas the main goal of this project is to retain that data to
aid optimisation and evaluation of Postfix restrictions.  \gls{LMA} fails
to parse Postfix log files generated on Solaris hosts because the fields
automatically prepended to each log line differ from those added on Linux
hosts; log files from Solaris hosts (and possibly other operating systems)
thus require preprocessing before parsing by \gls{LMA}.  Once the
preprocessing has been performed on the \numberOFlogFILES{} test log files
\gls{LMA} parses the log files without complaint, although it produced 32
entries in its output file for every rejection in the input log file; it
also missed some 40\% of delivered mail.  Once these glaring deficiencies
were discovered the author did not waste any more time checking the
results.

\gls{LMA} does provide some simple reports: IP-STORY, FROM-STORY,
DAILY-EMAIL and DAILY-REJECT\@.  The first three extract \gls{CSV} lines
for the specified client \gls{IP} address, sender address, or date
respectively.  DAILY-REJECT initially failed with an error message (XXX ADD
THE ERROR MESSAGE) from the Perl interpreter; after making some corrections
to the code it worked, and extracted the \gls{CSV} lines for the specified
day where the \gls{SMTP} code signifies a rejection.  All of these reports
are trivially simple to produce from the \gls{CSV} file using the standard
Unix tool \texttt{awk}\glsadd{awk}; the most complicated, DAILY-REJECT, is
merely:

@@

@@

XXX THE BLURB MENTIONS \gls{SQL} DB\@; CAN THAT BE ACCESSED SEPARATELY, IS
THE FORMAT DOCUMENTED, ETC\@?  XXX EXPERIMENT WITH THE REPORTS IN THE DEMO
VERSION\@.  XXX EVERYTHING FROM HERE DOWN NEEDS TO BE CHECKED AND/OR
REWRITTEN\@.

The web interface allows creation of dynamic reports based on any field,
but because of the associated cost the author has not experimented with it.
As the company charges per report the user wishes to generate, the data
store (if there is one) is deliberately inaccessible and undocumented to
prevent the user bypassing the program and generating their own reports.

The source code is available only in an obfuscated form (presumably for a
fee), and the product is quite expensive, requiring a \euros{100} + VAT
license per report to be generated; in contrast \parsername{} is free and
the code is freely available.

Sawmill's web interface supports searching on any combination of the fields
it extracts, but the interface is neither as simple to use nor as
informative as the interface provided by Splunk.  The administrative
interface is easy to use --- it took only five minutes to start parsing a
directory of log files.

@@

Splunk supports reading compressed log files without any configuration by
the user.  The free version of Splunk limits the volume of data indexed per
day to 500MB, though a trial Enterprise licence is available that allows
indexing of up to 5GB of data per day.  In 2007 the cheapest licensed
version cost \$5000 plus \$1000 support, and limited the volume of data
indexed per day to 500MB\@.  Prices were removed from the Splunk website
during 2008; now Splunk's sales team must be contacted for a quote.
Typical log file sizes for a small scale mail server are given in
\sectionref{parser efficiency}.

@@

\begin{table}[htb]
    \caption{Summary of parsers' features}
    \empty{}\label{Summary of parsers' features}
    \begin{tabular}{llllll}
        Parser          & Parsed test   & Data              & Custom            & Documentation  & Source \\
                        & log files?    & store?            & reports?          & quality?       & code?  \\
        \tableline{}%
        \gls{LMA}          & No            & Yes               & No                & Poor           & Yes    \\ 
        Pflogsumm       & Yes           & No                & Partial \dag{}    & Good           & Yes    \\
        Sawmill         & Yes           & Yes               & Searches          & Reasonable     & No     \\
        Splunk          & Yes           & Yes               & Searches          & Abundant       & No     \\
                        &               &                   & \& reports        & but poor       &        \\
        Isoqlog         & No            & Yes               & No                & Poor           & Yes    \\
        AWStats         & Partially     & Yes               & Partial \dag{}    & Good           & Yes    \\
        Anteater        & No            & No                & No                & Poor           & Yes    \\
        YAALA           & No            & Yes \ddag{}       & Searches          & Poor           & Yes    \\
        Lire            & Yes           & Yes               & Yes               & Reasonable     & Yes    \\
        Logrep          & No            & Yes               & No                & Poor           & Yes    \\
        \parsername{}   & Yes           & Yes \nialpha{}    & No \nibeta{}      & XXX            & Yes    \\
        \tableline{}
    \end{tabular}

@@

        \item [\nialpha{}] \parsername{} is the only parser with
            documentation for its data store.

        \item [\nibeta{}] \parsername{} defers report generation to
            subsequent programs, but all the necessary data and
            documentation to produce reports is provided.

@@

>>>>

\newpage{} %XXX

@@

It is important to compare and contrast newly developed programs,
algorithms, and parsers against those already available, to accurately
judge what improvements, if any, are delivered by the newcomers.  Ten other
parsers have been reviewed as part of the background research for this
project, starting with the only previously published research in this area
that the author is aware of.  There are not that many previously developed
Postfix log file parsers --- indeed it was quite difficult to find ten
parsers to review for this project --- and the functionality offered ranges
from quite basic to much more mature, depending on the needs of the author
of the parser.  The programs are not reviewed from an independent
viewpoint; the objective is to compare and contrast each program with
\parsername{}.  There are some important differences between \parsername{}
and the parsers reviewed here:

@@

    \item None of the parsers reviewed perform the kind of advanced parsing
        and correlation required for this project (described in
        \sectionref{complications})

@@

    \item The parsers reviewed all produce a report of varying complexity
        and detail, whereas \gls{PLP} is designed to be a generic parser,
        extracting data and leaving generation of reports from the data to
        other programs.  Using an \gls{SQL} database simplifies the process
        of generating such reports (discussed in \sectionref{database as
        API}); some simple examples are given in \sectionref{motivation}.

        The parser developed for this project is designed to enable much
        more detailed log file analysis by providing a stable platform for
        subsequent programs to develop upon.

@@

    \item A minor difference is that most parsers do not handle compressed
        files; both \parsername{} and Splunk handle them transparently,
        without user intervention; Sawmill and Lire can be configured to
        support compressed files, but Sawmill exhibits a dramatic increase
        in parsing time when doing so.  Although this is a minor
        difference, support for reading compressed log files is quite
        helpful, as it dramatically reduces the disk space required to
        store old log files.

    \item Some of the parsers reviewed save the data extracted to a data
        store, but the majority discard all data once they have finished
        generating the report, making historical analysis impossible.

@@

Each of the parsers was tested with the \numberOFlogFILES{} test log files
described in \sectionref{parser efficiency}.  The data extracted by
\parsername{} is documented in \sectionref{connections table} and
\sectionref{results table}.

XXX LIST THE DATA NOT EXTRACTED BY EACH PARSER\@.

@@

@@

Very little documentation is provided with \gls{LMA}, though some
documentation is available in~\cite{log-mail-analyser}.  Studying the
source code is informative, though this author had difficulty as the
authors of \gls{LMA} wrote in Italian.

@@

The \gls{CSV} record format is not documented, but the output file contains
a comment giving the format:\newline{} \texttt{\# Timestamp|Nome Client|IP
Client|IP Server|From|To|Status|Size} \newline{}\gls{LMA} treats lines
starting with \texttt{\#} as comments, but not all \gls{CSV} parsers will,
as there is no standard for comments in \gls{CSV} files.

@@

\gls{LMA} does not deal with any of the complications discussed in
\sectionref{complications}.  It does not differentiate between different
types of rejections, so it is not suitable for the purposes of this
project; the data about which restriction caused the rejection is
discarded, whereas the main goal of this project is to retain that data to
aid optimisation and evaluation of Postfix restrictions.  \gls{LMA} fails
to parse Postfix log files generated on Solaris hosts because the fields
automatically prepended to each log line differ from those added on Linux
hosts; log files from Solaris hosts (and possibly other operating systems)
thus require preprocessing before parsing by \gls{LMA}.  Once the
preprocessing has been performed on the \numberOFlogFILES{} test log files
\gls{LMA} parses the log files without complaint, although it produced 32
entries in its output file for every rejection in the input log file; it
also missed some 40\% of delivered mail.  Once these glaring deficiencies
were discovered the author did not waste any more time checking the
results.

\gls{LMA} does provide some simple reports: IP-STORY, FROM-STORY,
DAILY-EMAIL and DAILY-REJECT\@.  The first three extract \gls{CSV} lines
for the specified client \gls{IP} address, sender address, or date
respectively.  DAILY-REJECT initially failed with an error message from the
Perl interpreter;\footnote{The error message was:
\newline{}\texttt{Undefined subroutine \&main::LIST called at queryDB.pl
line 372.}} after making some corrections to the code it worked, and
extracted the \gls{CSV} lines for the specified day where the \gls{SMTP}
code signifies a rejection.  All of these reports are trivially simple to
produce from the \gls{CSV} file using the standard Unix tool
\texttt{awk}\glsadd{awk}; the most complicated, DAILY-REJECT, is merely:

@@

The source code is available in an obfuscated form, and the product is
quite expensive, requiring a \euros{100} + VAT licence per report
(discounts are available when buying multiple licences); in contrast
\parsername{} is free and the code is freely available.  Sawmill is
supplied with thorough and well written documentation; everything the
author searched for was documented, except the MySQL database schema.  A
commercial version of MySQL is required due to MySQL licensing
restrictions, but Sawmill's documentation explains why and includes
instructions on how to compile Sawmill so that it can use a non-commercial
version of MySQL\@.

Sawmill's web interface supports searching on any combination of the fields
it extracts, but the interface is neither as simple to use nor as
informative as the interface provided by Splunk.  The administrative
interface is much easier to use than Splunk's: it took only five minutes to
start parsing a directory of log files.  Searches performed using the web
interface produced reasonable results, which were not verified in detail.

@@

@@

Splunk supports reading compressed log files without any configuration by
the user.  The free version of Splunk limits the volume of data indexed per
day to 500MB, though a trial Enterprise licence is available that allows
indexing of up to 5GB of data per day.  In 2007 the cheapest licenced
version cost \$5000 plus \$1000 support, and limited the volume of data
indexed per day to 500MB\@.  Prices were removed from the Splunk website
during 2008; now Splunk's sales team must be contacted for a quote.
Typical log file sizes for a small scale mail server are given in
\sectionref{parser efficiency}.

@@

\begin{table}[htb]
    \caption{Summary of parsers' features}
    \empty{}\label{Summary of parsers' features}
    \begin{tabular}{llllll}
        Parser          & Parsed test   & Data              & Custom            & Documentation  & Source       \\
                        & log files?    & store?            & reports?          & quality?       & code?        \\
        \tableline{}%
        \gls{LMA}       & No            & Yes               & No                & Poor           & Yes          \\ 
        Pflogsumm       & Yes           & No                & Partial \dag{}    & Good           & Yes          \\
        Sawmill         & Yes           & Yes               & Searches          & Very good      & \nialpha{}   \\
        Splunk          & Yes           & Yes               & Searches          & Abundant       & No           \\
                        &               &                   & \& reports        & but poor       &              \\
        Isoqlog         & No            & Yes               & No                & Poor           & Yes          \\
        AWStats         & Partially     & Yes               & Partial \dag{}    & Good           & Yes          \\
        Anteater        & No            & No                & No                & Poor           & Yes          \\
        YAALA           & No            & Yes \ddag{}       & Searches          & Poor           & Yes          \\
        Lire            & Yes           & Yes               & Yes               & Reasonable     & Yes          \\
        Logrep          & No            & Yes               & No                & Poor           & Yes          \\
        \parsername{}   & Yes           & Yes \nibeta{}     & No \nichi{}       & XXX            & Yes          \\
        \tableline{}
    \end{tabular}

@@

        \item [\nialpha{}] Sawmill's source code is available in an
            obfuscated form, so that customers can compile it on platforms
            that pre-compiled binaries are not available for.

        \item [\nibeta{}] \parsername{} is the only parser with
            documentation for its data store.

        \item [\nichi{}] \parsername{} defers report generation to
            subsequent programs, but all the necessary data and
            documentation to produce reports is provided.

@@

\clearpage{}
